The onset of hydrodynamically driven waving of marine grass in flow

### 2014-09-26 Paper/grass.tex
Post Mahesh's corrections
- Fix real part vs Reynolds number conflict
- Remove introduction of Re for blade scale Reynolds number
- Minor cosmetic changes

$P$ the pressure, $\mu$ the turbulent viscosity and $\mathbf{g}$ the acceleration due to gravity. 
The Reynolds number (\Rey) of the flow based on the scale of a grass blade is expected to be $O(10^2-10^3)$, and hence the drag force is dominated by pressure and modeled as 

@@

We solve this eigenvalue problem and a threshold in $\Rey$ of $O(100)$, above which the flow is unstable (Re($\sigma$)$>$0), emerges from the 
solution of eq.~\eqref{Orr-somerfield}. The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown 
in fig.~\ref{Re_vs_delta}, and is found to compare well with experimental observations \cite{Ghisal02}.
The threshold Reynolds number increases with the grass density, indicating a competition between the 
destabilizing shear in the flow, and the stabilizing effect of damping due to vegetation drag.
Previous linear stability calculations for terrestrial grass do not exhibit a threshold flow condition 
either due to the exclusion of the vegetation drag in their model \cite{Raupach96}, or because of ad-hoc
assumptions for the mean velocity profile \cite{Raupach96,Delangre06}.

@@

The distinct asymptotic behavior of the two modes as $\Ndg \gg 1$ allows us to understand the mechanism of the instability. 
Mode 1 asymptotically localizes to the boundary layer near the grass tip, and exhibits an asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim (H/\delta)$ (or $\Rey \propto {\Ndg}^{1/2}$) at the threshold. 
This limit can be understood through asymptotically estimating the sizes of the terms in eq.~\eqref{Orr-somerfield} using $D\sim \delta^{-1}$, $\sigma \sim \delta^{-1}$, and $U\sim \delta$; the magnitude of the advection term is $\Rey/\delta^3 \sim \Rey^2 \Ndg $, and the viscous term (or the vegetation drag term, which is found to scale identically) in the boundary layer is $\delta^{-4} \sim (\Rey \Ndg)^{4/3}$. 
The terms balance when $\Rey \sim {\Ndg}^{1/2}$ (or $\Rey \sim H/\delta$) leading to a simplification of eq.~\eqref{Orr-somerfield} as
\begin{align}
(D^2-k^2)^2 \phi- 2 \Rey \Ndg D( SUD\phi) = \Rey \sigma (D^2-k^2)\phi,
\label{eqn:mode1asymp}
\end{align} 
in a region of thickness O($\delta$) near $y=\hg$.
Since the relative magnitude of the terms on the r.h.s. to those on the l.h.s. of this equation is $\Rey/\Ndg^{1/2}$, we expect identical asymptotic behavior for fixed $\Rey/\Ndg^{1/2}$. Therefore the threshold obtained for Mode 1 is $\Rey \propto \Ndg^{1/2}$ (or $\Rey \propto H/\delta$) explaining the numerically observed asymptote (see fig.~\ref{Re_vs_delta}). 
This analysis also concludes that the mode structure is self-similar over the length scale $\delta$ for fixed $\Rey/\Ndg^{1/2}$; the verification of this expectation is shown in fig.~\ref{Asymptotic_mode}, supporting this argument.

@@

\acknowledgments
MMB was hosted by Brown University during this work and was supported by the OIST Graduate University with subsidy funding from the Cabinet Office, Japan.

>>>>

$P$ the pressure, $\mu$ the eddy viscosity (the flow is usually turbulent) and $\mathbf{g}$ the acceleration due to gravity. 
The Reynolds number of the flow based on the scale of a grass blade is expected to be $O(10^2-10^3)$, and hence the drag force is dominated by pressure and modeled as 

@@

We solve this eigenvalue problem and a threshold in $\Rey$ of $O(100)$, above which the flow is unstable (real part of $\sigma$ is positive), emerges from the 
solution of eq.~\eqref{Orr-somerfield}. The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown 
in fig.~\ref{Re_vs_delta}, and is found to compare well with experimental observations \cite{Ghisal02}.
The threshold Reynolds number increases with the grass density, indicating a competition between the 
destabilizing shear in the flow, and the stabilizing effect of damping due to vegetation drag.
Previous linear stability calculations for terrestrial grass do not exhibit a threshold flow condition 
either due to the exclusion of the vegetation drag in their model \cite{Raupach96}, or because of ad-hoc
assumptions for the mean velocity profile \cite{Raupach96,Delangre06}.

@@

The distinct asymptotic behavior of the two modes as $\Ndg \gg 1$ allows us to understand the mechanism of the instability. 
Mode 1 asymptotically localizes to the boundary layer near the grass tip, and exhibits an asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim (H/\delta)$ (or $\Rey \propto {\Ndg}^{1/2}$) at the threshold. 
This limit can be understood through asymptotically estimating the sizes of the terms in eq.~\eqref{Orr-somerfield} using $D\sim \delta^{-1}$, $\sigma \sim \delta^{-1}$, and $U\sim \delta$; the magnitude of the advection term is $\Rey/\delta^3 \sim \Rey^2 \Ndg $, and the viscous term (or the vegetation drag term, which is found to scale identically) in the boundary layer is $\delta^{-4} \sim (\Rey \Ndg)^{4/3}$. 
The terms balance when $\Rey \sim {\Ndg}^{1/2}$ (or $\Rey \sim H/\delta$) leading to a simplification of eq.~\eqref{Orr-somerfield} as
\begin{align}
(D^2-k^2)^2 \phi- 2 \Rey \Ndg D( SUD\phi) = \sigma \Rey (D^2-k^2)\phi,
\label{eqn:mode1asymp}
\end{align} 
in a region of thickness O($\delta$) near $y=\hg$.
Since the relative magnitude of the terms on the r.h.s. to those on the l.h.s. of this equation is $\Rey/\Ndg^{1/2}$, we expect identical asymptotic behavior for fixed $\Rey/\Ndg^{1/2}$. Therefore the threshold obtained for Mode 1 is $\Rey \propto \Ndg^{1/2}$ (or $\Rey \propto H/\delta$) explaining the numerically observed asymptote (see fig.~\ref{Re_vs_delta}). 
This analysis also concludes that the mode structure is self-similar over the length scale $\delta$ for fixed $\Rey/\Ndg^{1/2}$; the verification of this expectation is shown in fig.~\ref{Asymptotic_mode}, supporting this argument.

@@

\acknowledgments
MMB was hosted by Brown University during this work and was supported by the OIST Graduate University with subsidy funding from the Cabinet Office, Japan. We thanks Heidi Nepf and Marco Ghisalberti for helpful discussions.
